\chapter{Comparative Analysis}
\label{chap:comparison}
This chapter provides a cross-sample comparison of the three Malbus APK variants analysed in the previous chapters: \textbf{Malbus \oner \ }, \textbf{Malbus \twor \ }, and \textbf{Malbus \threer \ }.
The goal is to identify common patterns, progressive evolution of malicious capabilities, and differences in evasion sophistication across the family.

% ─────────────────────────────────────────────────────────────────────────────
\section{VirusTotal Detection}
% ─────────────────────────────────────────────────────────────────────────────

All three samples were submitted to VirusTotal and flagged as malicious by a comparable number of security engines.
The suggested threat label assigned by VirusTotal is identical across all samples --- \texttt{trojan.andr/apkdownloader} --- and the Tencent engine correctly identified each sample by referencing its specific APK package name (\texttt{a.privacy.MalBuslmh}).

\begin{table}[H]
\centering
\begin{tabular}{|l|c|c|c|}
\hline
\textbf{Metric} & \textbf{\oner \ } & \textbf{\twor \ } & \textbf{\threer \ } \\
\hline
Malicious detections & 40/75 & 40/75 & 38/75 \\
Reputation score & $-1$ & $-2$ & $0$ \\
Suggested threat label & \multicolumn{3}{c|}{\texttt{trojan.andr/apkdownloader}} \\
Dominant engine category & trojan (18) & trojan (17) & trojan (17) \\
Secondary engine category & downloader (5) & downloader (4) + dropper (1) & downloader (5) \\
Total submissions & 377 & 354 & -- \\
Alternative filenames & Yes & Yes & Yes \\
\hline
\end{tabular}
\caption{VirusTotal detection summary across the three APK samples}
\end{table}

Notably, well-known vendors such as Microsoft Defender, McAfee, and Malwarebytes failed to detect any of the three samples as malicious, highlighting a significant gap in coverage for this malware family.

% ─────────────────────────────────────────────────────────────────────────────
\section{Permissions and Attack Surface}
\label{sec:permissions}
% ─────────────────────────────────────────────────────────────────────────────

All three APKs share a core permission set consistent with the same malware family.
The most security-relevant permissions present across all samples are:

\begin{itemize}
    \item \texttt{ACCESS\_FINE\_LOCATION} and \texttt{ACCESS\_COARSE\_LOCATION} --- device tracking via GPS and cell towers.
    \item \texttt{READ\_EXTERNAL\_STORAGE} / \texttt{WRITE\_EXTERNAL\_STORAGE} --- access to user files on external memory.
    \item \texttt{REQUEST\_INSTALL\_PACKAGES} --- silent installation of additional APKs, a key dropper indicator.
    \item \texttt{INTERNET} --- unrestricted network access, used for remote configuration and payload delivery.
    \item GCM/FCM push messaging (\texttt{C2DMESSAGE}, \texttt{C2DM.RECEIVE}) --- allows remote triggering of behaviour.
\end{itemize}

Two per-version differences emerge when comparing the manifests.
First, \texttt{WRITE\_HISTORY\_BOOKMARKS} and \texttt{READ\_GSERVICES} are present in \oner \  and \threer \ , but are absent in \twor \ .
The \texttt{WRITE\_HISTORY\_BOOKMARKS} permission grants read and write access to the Android stock browser's history and bookmarks. Its presence in two out of three variants is not consistent with a legitimate bus-information application and suggests either browser history exfiltration for user profiling --- a capability observed in banking-style trojans that fingerprint visited financial sites --- or click-fraud redirection, in which the malware inserts attacker-controlled bookmarks to inflate advertising revenues. The permission's absence in \twor \  may reflect an intentional reduction of the detectable attack surface for that release, subsequently reversed in \threer \ .
Second, \threer \  uniquely declares \texttt{CALL\_PHONE}, \texttt{RECORD\_AUDIO}, and \texttt{READ\_PHONE\_STATE} in its manifest, permissions entirely absent from \oner \  and \twor \ .
Combined with the Crashlytics SDK declared only in \threer \ , this indicates a measurably expanded attack surface in the latest variant.

% ─────────────────────────────────────────────────────────────────────────────
\section{Behavioural Tags and Evasion Capabilities}
% ─────────────────────────────────────────────────────────────────────────────

A clear evolution in anti-analysis sophistication is observable across the three versions.
While \oner \  employs obfuscation and Java Reflection, versions \twor \  and \threer \  introduce explicit VM and debugger evasion mechanisms.
Noteworthy is that \threer \  only carries the three anti-analysis tags and lacks the \texttt{obfuscated}, \texttt{reflection}, \texttt{telephony}, and \texttt{checks-gps} tags visible in \oner \  and \twor \ ; this likely reflects a difference in VirusTotal sandbox coverage for this sample rather than the removal of those capabilities from the code.

\begin{table}[H]
\centering
\begin{tabular}{|l|c|c|c|}
\hline
\textbf{Behavioural Tag} & \textbf{\oner \ } & \textbf{\twor \ } & \textbf{\threer \ } \\
\hline
obfuscated & \checkmark & \checkmark & $\times$ \\
reflection & \checkmark & \checkmark & $\times$ \\
checks-gps & \checkmark & \checkmark & $\times$ \\
telephony & \checkmark & \checkmark & $\times$ \\
detect-debug-environment & $\times$ & \checkmark & \checkmark \\
checks-cpu-name & $\times$ & \checkmark & \checkmark \\
sets-process-name & $\times$ & \checkmark & \checkmark \\
\hline
\end{tabular}
\caption{VirusTotal behavioural tags across the three APK samples}
\end{table}

This progression confirms that the malware author actively improved the evasion layer between \oner \  and the later variants, adding sandbox detection and process name manipulation to harden the payload against dynamic analysis environments.

% ─────────────────────────────────────────────────────────────────────────────
\section{Static Analysis: Common Architecture}
% ─────────────────────────────────────────────────────────────────────────────

The static analysis of all three samples reveals a shared internal architecture, confirming they are regional variants of the same malware family.
The core classes --- \texttt{MainActivity}, \texttt{BaseMainActivity}, \texttt{EnvConstant}, \texttt{BusApiWeb}, \texttt{AppInfoActivity} --- are present and functionally equivalent across all three APKs, differing only in regional bus endpoint URLs and package names (\texttt{cwbus}, \texttt{gjbus}, \texttt{jjbus}).

The following structural elements are common to all three samples:

\begin{itemize}
    \item Loading of a native library via \texttt{System.loadLibrary("Audio3.0")} in \texttt{BaseMainActivity}.
    \item Native methods \texttt{startUpdate()} and \texttt{updateApplication()}, consistent with dropper-like update logic.
    \item Remote environment loading mechanism (\texttt{getEnvironmentProperties(\ldots)}) that allows runtime behaviour modification without a new APK release.
    \item Hardcoded advertising and analytics keys in \texttt{EnvConstant}: AdMob, AdLib, Facebook, Google Analytics (UA-56668425-6).
    \item Unencrypted HTTP communication in \texttt{BusApiWeb}, exposing bus-stop data to man-in-the-middle interception.
    \item A cosmetic \texttt{AppInfoActivity} displaying a legitimate regional bus URL to disguise the application's true nature.
\end{itemize}

An SDK-level difference is worth noting: \threer \  additionally includes the \textbf{Crashlytics} crash-reporting SDK and references \textbf{Firebase} more prominently in \texttt{MainActivity}, while \oner \  and \twor \  rely on AdFit (Kakao) without Crashlytics.

% ─────────────────────────────────────────────────────────────────────────────
\section{C2 Infrastructure Status}
\label{sec:c2status}
% ─────────────────────────────────────────────────────────────────────────────

As part of the dynamic analysis, the primary C2 endpoint shared by all three variants,\\
\texttt{busappmanager.elasticbeanstalk.com}, was directly probed via \texttt{curl} requests to the known API paths (\texttt{/v1/env/get.json}, \texttt{/v1/category/list.json}, \texttt{/v1/issue/my.json}).
The server returned no response on any path: the host did not resolve to an active IP address, confirming that the AWS Elastic Beanstalk environment hosting the Malbus distribution infrastructure has been decommissioned.

This finding has two important implications.
First, the dropper mechanism documented in \texttt{BaseMainActivity} --- remote environment loading via \texttt{getEnvironmentProperties(\ldots)}, followed by payload download and silent installation --- could not be triggered end-to-end during dynamic analysis, because the C2 backend is no longer operational.
The infection chain stages~2 through~4 described in the McAfee Malbus report (payload request, delivery, and installation of the second-stage trojan) are therefore confirmed at the structural and static level only; live execution was not possible.
Second, the C2 inactivity explains the database download errors observed in Genymotion across all three samples: since the remote environment configuration cannot be retrieved, the application is unable to fetch its bus database or payload URL, and fails with a network error on startup.

The \texttt{smartapps.cafe24.com} variant-specific endpoints (\texttt{/gjbus\_android/environment2.json}, \texttt{/cwbus\_android/}, \texttt{/jjbus\_android/environment.json}) were independently verified as unreachable at the time of analysis (June 2026), confirming that the entire Malbus remote configuration infrastructure is defunct.

% ─────────────────────────────────────────────────────────────────────────────
\section{MobSF Static Analysis Comparison}
% ─────────────────────────────────────────────────────────────────────────────

MobSF confirms consistent and severe security issues in all three samples.
All APKs share the same certificate signing profile: SHA-1 with RSA, v1 scheme only, issued by \texttt{CN=Myoung-Hun Lee} --- making all three vulnerable to the Janus attack on Android 5.0--8.0.
Notably, \threer \  received an explicit MobSF risk score of \textbf{38/100 (Grade C, HIGH RISK)}, encompassing 8 privacy trackers detected out of 432 checked and 4 critical permission hotspots; the other two samples were not assigned a numerical score in the analysed report.

\subsection{High-Severity Findings}

\begin{table}[H]
\centering
\renewcommand{\arraystretch}{1.15}
\begin{tabular}{|l|c|c|c|}
\hline
\textbf{Finding} & \textbf{\oner \ } & \textbf{\twor \ } & \textbf{\threer \ } \\
\hline
Janus vulnerability & \checkmark & \checkmark & \checkmark \\
SHA-1 weak hashing (certificate) & \checkmark & \checkmark & \checkmark \\
Low minSdk = 14 & \checkmark & \checkmark & \checkmark \\
Task Hijacking \& StrandHogg & \checkmark & \checkmark & StrandHogg 2.0 \\
Debug configuration enabled & \checkmark & \checkmark & \checkmark \\
Weak encryption in AdLib library & \checkmark & \checkmark & $\times$ \\
SharedPreferences & \checkmark & \checkmark & \checkmark \\
Privacy trackers ($>$8) & \checkmark & \checkmark & \checkmark \\
Logging sensitive info\tablefootnote{Classified as a LOW-severity (informational) finding in Chapters~1 and~2; elevated to HIGH severity by MobSF in Chapter~3 for \threer \ .} & $\times$ & $\times$ & \checkmark \\
Insecure random number generator & $\times$ & $\times$ & \checkmark \\
Weak cryptographic primitive & $\times$ & $\times$ & \checkmark \\
libAudio3.0: no stack canary & $\times$ & $\times$ & \checkmark \\
\hline
\textit{Total HIGH findings} & \textit{9} & \textit{9} & \textit{10} \\
\hline
\end{tabular}
\caption{MobSF high-severity findings across the three APK samples}
\label{tab:mobsf-high}
\end{table}

The total count of high-severity findings is 9 for \oner \  and \twor \ , and rises to \textbf{10} for \threer \ .
The composition shifts as follows: the weak AdLib encryption finding (absent in \threer \ ) is replaced by four code-level issues --- logging of sensitive information (elevated from LOW in earlier variants), insecure RNG, weak MD5 cryptography, and the unprotected native library --- while the full Task Hijacking weakness is reduced to StrandHogg~2.0 only.

\subsection{Medium-Severity Findings}

\begin{table}[H]
\centering
\renewcommand{\arraystretch}{1.15}
\begin{tabular}{|l|c|c|c|}
\hline
\textbf{Finding} & \textbf{\oner \ } & \textbf{\twor \ } & \textbf{\threer \ } \\
\hline
\texttt{android:allowBackup=true} & \checkmark & \checkmark & \checkmark \\
Exported / unprotected components & \checkmark & \checkmark & \checkmark \\
INSTALL\_PACKAGES external permission & \checkmark & \checkmark & \checkmark \\
Read/write external storage & \checkmark & \checkmark & \checkmark \\
Hardcoded secrets (API keys) & \checkmark & \checkmark & \checkmark \\
Insecure WebView usage & \checkmark & \checkmark & \checkmark \\
Insecure RNG\tablefootnote{Also flagged as HIGH severity for \threer \  (see Table~\ref{tab:mobsf-high}); MobSF reports it at both severity levels in Chapter~3.} & \checkmark & \checkmark & \checkmark \\
MD5 weak hashing\tablefootnote{Also flagged as HIGH severity for \threer \  (see Table~\ref{tab:mobsf-high}); MobSF reports it at both severity levels in Chapter~3.} & \checkmark & \checkmark & \checkmark \\
SQL injection via raw queries & \checkmark & \checkmark & \checkmark \\
\texttt{CALL\_PHONE} / \texttt{RECORD\_AUDIO} permissions & $\times$ & $\times$ & \checkmark \\
\hline
\end{tabular}
\caption{MobSF medium-severity findings across the three APK samples}
\end{table}

For \threer \ , MobSF reports Insecure RNG and MD5 at \emph{both} medium and high severity, reflecting their dual presence across the scorecard sections of Chapter~3.
The unique presence of \texttt{CALL\_PHONE}, \texttt{RECORD\_AUDIO}, and \texttt{READ\_PHONE\_STATE} in \threer \  was also flagged as a medium-severity concern, consistent with the expanded manifest permissions discussed in Section~\ref{sec:permissions}.

% ─────────────────────────────────────────────────────────────────────────────
\section{YARA Rule Matches}
% ─────────────────────────────────────────────────────────────────────────────

All three samples matched the \texttt{IcedID} structural YARA rule from the \texttt{CAPEv2} repository (author: \texttt{kevoreilly/threathive}).
Although IcedID is historically associated with a Windows banking trojan, the matched rule targets dropper and loader architectures, not the specific malware family.
This consistent structural match across all three variants reinforces the hypothesis that the Malbus family was deliberately engineered with an advanced payload-delivery architecture designed to withstand analysis and deliver second-stage payloads on demand.

% ─────────────────────────────────────────────────────────────────────────────
\section{Overall Family Assessment}
% ─────────────────────────────────────────────────────────────────────────────

The comparative analysis confirms that Malbus \oner \ , \twor \ , and \threer \  are three regional variants of the same trojanised application family, sharing an identical malicious core while adapting the legitimate-facing functionality to different South Korean cities (Changwon, Gwangju, Jeonju).

The evolution across versions is twofold.
First, anti-analysis and VM evasion capabilities are absent in \oner \  but fully present in \twor \  and \threer \  (\texttt{detect-debug-environment}, \texttt{checks-cpu-name}, \texttt{sets-process-name}).
Second, \threer \  introduces additional hardware-level permissions (\texttt{CALL\_PHONE}, \texttt{RECORD\_AUDIO}, \texttt{READ\_PHONE\_STATE}), the Crashlytics SDK, and a native library (\texttt{libAudio3.0.so}) compiled without stack canary or fortify function protections --- all absent in the earlier samples.

In all other dimensions --- certificate signing, dropper architecture, remote configuration, advertising SDK integration, and the core MobSF finding set --- the three samples are functionally identical, indicating a mature and deliberately maintained malware family rather than opportunistic or independent variants.
The confirmed inactivity of the shared C2 infrastructure (see Section~\ref{sec:c2status}) prevented end-to-end reproduction of the second-stage payload delivery; the dropper mechanism is nonetheless fully evidenced at the structural and static analysis level across all three variants.
