\chapter*{References}
\addcontentsline{toc}{chapter}{References} % Add References to Table of Contents
\label{sec:references}
\listoftodos[Todos]
\listoffigures
\begin{thebibliography}{99}

    % ── Course material ──────────────────────────────────────────────────────
    \bibitem{Lab00} Dependability Laboratory Lecture Notes: Introduction.
    University of Pisa, 2024--2025.

    \bibitem{Lab01} Dependability Laboratory Lecture Notes: Malware Analysis 1.
    University of Pisa, 2024--2025.

    \bibitem{Lab02} Dependability Laboratory Lecture Notes: Malware Analysis 2.
    University of Pisa, 2024--2025.

    \bibitem{Lab03} Dependability Laboratory Lecture Notes: Malware Analysis 3.
    University of Pisa, 2024--2025.

    \bibitem{Lab04} Dependability Laboratory Lecture Notes: Malware Analysis 4.
    University of Pisa, 2024--2025.

    % ── Malware family documentation ─────────────────────────────────────────
    \bibitem{McAfee-Malbus} C.\ Beek and S.\ Rubin,
    ``Malbus --- Trojanized Korean Bus App on Google Play,''
    McAfee Mobile Research, 2019.
    \url{https://www.mcafee.com/blogs/other-blogs/mcafee-labs/malbus-malware-takes-the-bus-route-to-android-users/}

    % ── Analysis tools ───────────────────────────────────────────────────────
    \bibitem{VirusTotal} VirusTotal --- Analyze suspicious files, domains, IPs
    and URLs to detect malware and other breaches.
    \url{https://www.virustotal.com/}

    \bibitem{MobSF} Mobile Security Framework (MobSF) --- An automated,
    all-in-one mobile application (Android/iOS/Windows) pen-testing, malware
    analysis and security assessment framework.
    \url{https://github.com/MobSF/Mobile-Security-Framework-MobSF}

    \bibitem{JD-GUI} Java Decompiler (JD-GUI) --- A standalone graphical utility
    that displays Java sources from \texttt{.class} files.
    \url{https://java-decompiler.github.io/}

    \bibitem{Genymotion} Genymotion --- Android Virtual Device platform used as
    dynamic analysis environment.
    \url{https://www.genymotion.com/}

    % ── Vulnerability and security references ────────────────────────────────
    \bibitem{Janus} V.\ Apvrille, ``Janus: A new exploitation technique for
    Android,'' GuardSquare Research, 2017.
    \url{https://www.guardsquare.com/blog/new-android-vulnerability-allows-attackers-to-modify-apps-without-affecting-their-signatures}

    \bibitem{StrandHogg} Promon Security Research Team,
    ``StrandHogg 2.0 --- The 'evil twin' vulnerability,'' 2020.
    \url{https://promon.co/security-news/strandhogg-2/}

    \bibitem{CAPEv2-YARA} kevoreilly/threathive, ``IcedID YARA rule,''
    CAPEv2 ruleset on GitHub.
    \url{https://github.com/kevoreilly/CAPEv2}

    % ── Android platform ─────────────────────────────────────────────────────
    \bibitem{AndroidManifest} Android Developers,
    ``App Manifest Overview.''
    \url{https://developer.android.com/guide/topics/manifest/manifest-intro}

    \bibitem{AndroidPermissions} Android Developers,
    ``Permissions on Android.''
    \url{https://developer.android.com/guide/topics/permissions/overview}

\end{thebibliography}
